---
title: 'Exploration of All The News Corpus (2013-2020)'
author: "Bobby McHardy"
date: "January 06, 2020"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---

```{r preppack, include=FALSE}
library("tibble")
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(corrplot)
library(stringr)
library(plotly)
library(Hmisc)
library(zoo)
options(scipen = 999)
rm(list = ls())
```

## Loading in the data

We will now load the CORD-19 `.csv` of our found tokens and their corresponding frequencies.

```{r dataload, warning=FALSE}
# COVID-19 average cases and deaths by state and county in week leading up to 2020 election
data_news <- read.csv("all-the-news-2-1-table.csv")
data_news$year.cont <- data_news$year + ((data_news$month - 1) / 12) + data_news$day / 365
data_news
#data_language$freq <- as.numeric(data_language$freq)
```

Let's check the correlation of pos, neg, neu sentiments between title and article.

```{r correlations_1, message=FALSE, warning=FALSE}
cor(data_news[4:9])
```
Compute weighted averages on sentiment

```{r sentiment_weightedavg, message=FALSE, warning=FALSE}
# compute weighted average of title sentiment (with direction)
data_news$title.sent <- ifelse((data_news$title.pos.sent + data_news$title.neg.sent + data_news$title.neu.sent) == 0, 0, (data_news$title.pos.sent - data_news$title.neg.sent) / (data_news$title.pos.sent + data_news$title.neg.sent + data_news$title.neu.sent))
# compute weighted average of article sentiment (with direction)
data_news$article.sent <- ifelse((data_news$article.pos.sent + data_news$article.neg.sent + data_news$article.neu.sent) == 0, 0, (data_news$article.pos.sent - data_news$article.neg.sent) / (data_news$article.pos.sent + data_news$article.neg.sent + data_news$article.neu.sent))
# remove all the exact zero-weight sentiment (could be due to N/A content)
data_news <- data_news %>% filter(!(title.sent == 0 && article.sent == 0))
```

Now, we check correlations between weighted title and article sentiment.

```{r correlations_2, message=FALSE, warning=FALSE}
cor(data_news[13:14])
```

```{r viz_scatter_senticorr, message=FALSE, warning=FALSE}
# build chart
data_news %>% plot_ly(x = ~article.sent, y = ~title.sent,
                          type = "scatter") %>%
  layout(xaxis = list(range=c(-1,1)), yaxis = list(range=c(-1,1)))
```

Compute average of title and content.

```{r sentiment_weightedavg2, message=FALSE, warning=FALSE}
# compute weighted average of title sentiment (with direction)
data_news$sent <- data_news$title.sent + data_news$article.sent
```

Let's check our new synthetic sentiment measure.

```{r viz_histogram_sentiunidimensional, message=FALSE, warning=FALSE}
# build chart
data_news %>% plot_ly(x = ~sent,
                          type = "histogram") %>%
  layout(xaxis = list(range=c(-1,1)))
```

Display a timeline of sentiments, based on this new unidimensional sentiment model.

```{r viz_timeline_senti_aggregate1, message=FALSE, warning=FALSE}
# build chart
data_news %>% plot_ly(x = ~sent, y = ~year, color=~publication, type = "scatter",
                      text = paste0("<b>", data_news$publication,
                                    "</b><br><br>Article: <i>",
                                    data_news$title,
                                    "</i><br>URL: <a href='",
                                    data_news$url, "'>",
                                    substr(data_news$url, 0, 30),
                                    "...</a>"),
                      hovertemplate = "%{text}") %>%
  layout(xaxis = list(range=c(-1,1)), yaxis = list(range=c(2015,2021)))
```

```{r viz_timeline_senti_split1, message=FALSE, warning=FALSE}
# build chart
fig_cnn <- plot_ly(data_news %>% filter(publication == "CNN"),
                   x = ~sent, y = ~year, name = "CNN", type = "scatter")
fig_cnn <- fig_cnn %>% layout(xaxis = list(range=c(-1,1)), yaxis = list(range=c(2015,2021)))

fig_nyt <- plot_ly(data_news %>% filter(publication == "The New York Times"),
                   x = ~sent, y = ~year, name = "The New York Times", type = "scatter")
fig_nyt <- fig_nyt %>% layout(xaxis = list(range=c(-1,1)), yaxis = list(range=c(2015,2021)))

fig_fox <- plot_ly(data_news %>% filter(publication == "Fox News"),
                   x = ~sent, y = ~year, name = "Fox News", type = "scatter")
fig_fox <- fig_fox %>% layout(xaxis = list(range=c(-1,1)), yaxis = list(range=c(2015,2021)))

plotly::subplot(fig_cnn, fig_nyt, fig_fox)

rm(fig_cnn)
rm(fig_nyt)
rm(fig_fox)
```

```{r viz_timeline_senti_aggregate2, message=FALSE, warning=FALSE}
# build chart
fig <- plot_ly()
fig <- fig %>% add_trace(data = data_news %>% filter(year == 2016), x = ~sent, name = "2016", type = "violin")
fig <- fig %>% add_trace(data = data_news %>% filter(year == 2017), x = ~sent, name = "2017", type = "violin")
fig <- fig %>% add_trace(data = data_news %>% filter(year == 2018), x = ~sent, name = "2018", type = "violin")
fig <- fig %>% add_trace(data = data_news %>% filter(year == 2019), x = ~sent, name = "2019", type = "violin")
fig <- fig %>% add_trace(data = data_news %>% filter(year == 2020), x = ~sent, name = "2020", type = "violin")
fig <- fig %>% add_trace(data = data_news, x = ~sent, y = ~year, yaxis="y2", name = "Year") %>% 
  layout(yaxis2 = list(overlaying = "y", side = "right"))
fig
rm(fig)
```

```{r viz_timeline_senti_aggregate3, message=FALSE, warning=FALSE}
# build chart
data_news$sent.transform <- ((data_news$sent)/(sqrt((data_news$sent)^2))) * abs(data_news$sent)^(1/5)
data_news %>% plot_ly(x = ~year.cont, y = ~title.neg.sent, color=~publication, type = "scatter") %>%
  layout(yaxis = list(range=c(-.1, 0.6)))
#  layout(xaxis = list(range=c(-1,1)), yaxis = list(range=c(2015,2021)))
```

```{r viz_timeline_senti_aggregate4, message=FALSE, warning=FALSE}
# build sample data for display
data_news_sample <- data_news[sample(1:nrow(data_news), size = 100,
                                                 prob = (data_news$sent)^2, replace = F), ]
d_n_s_fox <- data_news_sample %>% filter(publication == "Fox News")
d_n_s_nyt <- data_news_sample %>% filter(publication == "The New York Times")
d_n_s_cnn <- data_news_sample %>% filter(publication == "CNN")
# build chart
fig <- plot_ly() %>%
  layout(title = list(text = "Popular Media Emotion Between 2016 and 2020"))
fig <- fig %>%
    add_trace(data = d_n_s_fox, color = I("#002885"),
                         x = ~sent, y = ~year.cont, yaxis = "y2", name = ~publication,
                         text = paste0("<b>", d_n_s_fox$publication,
                                    "</b><br><br>Article: <i>",
                                    d_n_s_fox$title,
                                    "</i><br>URL: <a href='",
                                    d_n_s_fox$url, "'>",
                                    substr(d_n_s_fox$url, 0, 30),
                                    "...</a>"),
                         hovertemplate = "%{text}") %>%
    add_trace(data = d_n_s_nyt, color = I("#555555"),
                         x = ~sent, y = ~year.cont, yaxis = "y2", name = ~publication,
                         text = paste0("<b>", d_n_s_nyt$publication,
                                    "</b><br><br>Article: <i>",
                                    d_n_s_nyt$title,
                                    "</i><br>URL: <a href='",
                                    d_n_s_nyt$url, "'>",
                                    substr(d_n_s_nyt$url, 0, 30),
                                    "...</a>"),
                         hovertemplate = "%{text}") %>%
    add_trace(data = d_n_s_cnn, color = I("#EC2029"),
                         x = ~sent, y = ~year.cont, yaxis = "y2", name = ~publication,
                         text = paste0("<b>", d_n_s_cnn$publication,
                                    "</b><br><br>Article: <i>",
                                    d_n_s_cnn$title,
                                    "</i><br>URL: <a href='",
                                    d_n_s_cnn$url, "'>",
                                    substr(d_n_s_cnn$url, 0, 30),
                                    "...</a>"),
                         hovertemplate = "%{text}") %>%
  layout(
    xaxis = list(title = "Article Emotion", tickmode = "array",
                 nticks = 2, tickvals = c(-1, 1),
                 ticktext = c("More Negative", "More Positive")),
    xaxis2 = list(title = "Left"),
    yaxis = list(title = "News Source"),
    yaxis2 = list(title = "Year", overlaying = "y", side = "right", autotick = F, dtick = 1),
    legend = list(x = 1.1)
  )
fig <- fig %>% add_trace(data = data_news %>% filter(publication == "Fox News"), color = I("#002885"),
                         x = ~sent, name = "Fox", type = "violin", hoveron="points", box = list(visible = T))
fig <- fig %>% add_trace(data = data_news %>% filter(publication == "The New York Times"), color = I("#555555"),
                         x = ~sent, name = "NYT", type = "violin", hoveron="points", box = list(visible = T))
fig <- fig %>% add_trace(data = data_news %>% filter(publication == "CNN"), color = I("#EC2029"),
                         x = ~sent, name = "CNN", type = "violin", hoveron="points", box = list(visible = T))
fig
rm(fig)
```
